Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs

### 2020-06-11 FDT2020/FDT.tex
comments asnwered
more changes

\onehalfspacing
\begin{document}{\bibliographystyle{alpha}}

\begin{abstract}
We present a new algorithm, Fractional Decomposition Tree (FDT) for finding a feasible solution for an integer program (IP) where all variables are binary. FDT runs in polynomial time and is guaranteed to find a feasible integer solution provided the integrality gap is bounded. The algorithm gives a construction for Carr and Vempala's theorem that any feasible solution to the IP's linear-programming relaxation, when scaled by the instance integrality gap, dominates a convex combination of feasible solutions. FDT is also a tool for studying the integrality gap of IP formulations.  We demonstrate that with experiments studying the integrality gap of two problems: optimally augmenting a tree to a 2-edge-connected graph and finding a minimum-cost 2-edge-connected multi-subgraph (2EC). We also give a simplified algorithm, Dom2IP, that more quickly determines if an instance has an unbounded integrality gap. We show that FDT's speed and approximation quality compare well to that of feasibility pump on moderate-sized instances of the vertex cover problem. For a particular set of hard-to-decompose fractional 2EC solutions, FDT always gave a better integer solution than the best previous approximation algorithm (Christofides).

\end{abstract}
\clearpage

@@

\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{abbrv}
\bibliography{haddadan-ref}

>>>>

\onehalfspacing

\title{Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs}

\author{\textsc{Robert Carr}\thanks{University of New Mexico  {\tt{bobcarr@unm.edu}}. This material is based upon research supported in part by
the U.S. Office of Naval Research under award number N00014-18-1-2099.}
\and \textsc{Arash Haddadan}\thanks{Carnegie Mellon University  {\tt{ahaddada@andrew.cmu.edu}}.}
\and \textsc{Cynthia A. Phillips}\thanks{Sandia National Laboratories  {\tt{caphill@sandia.gov}}.}}

\begin{document}{\bibliographystyle{alpha}}

\maketitle

\begin{abstract}
We present a new algorithm, Fractional Decomposition Tree (FDT) for finding a feasible solution for an integer program (IP) where all variables are binary. FDT runs in polynomial time and is guaranteed to find a feasible integer solution provided the integrality gap is bounded. The algorithm gives a construction for Carr and Vempala's theorem that any feasible solution to the IP's linear-programming relaxation, when scaled by the instance integrality gap, dominates a convex combination of feasible solutions. FDT is also a tool for studying the integrality gap of IP formulations.  We demonstrate that with experiments studying the integrality gap of two problems: optimally augmenting a tree to a 2-edge-connected graph and finding a minimum-cost 2-edge-connected multi-subgraph (2EC). We also give a simplified algorithm, Dom2IP, that more quickly determines if an instance has an unbounded integrality gap. We show that FDT's speed and approximation quality compare well to that of feasibility pump on moderate-sized instances of the vertex cover problem. For a particular set of hard-to-decompose fractional 2EC solutions, FDT always gave a better integer solution than the best previous approximation algorithm (Christofides).
\end{abstract}

@@

\bibliographystyle{abbrv}
\bibliography{FDT}

### 2020-06-11 FDT2020/Introduction.tex
more changes

This value depends on the constraints in (\ref{S}).  We cannot hope to find solutions for the IP with objective values better than $g(I)\cdot z_{LP}(I,c)$.

More generally we can define the integrality gap for a class of instances $\mathcal{I}$ as follows.% In this case, the integrality gap for problem $\mathcal{I}$ is
\begin{equation}\label{gapproblem}
g(\mathcal{I}) = \max_{c\geq 0 , I\in\mathcal{I}}\frac{z_{IP}(I,c)}{z_{LP}(I,c)}.
\end{equation}
For example, the aforementioned integrality gap of the subtour elimination relaxation for the 2EC is at most $\frac{3}{2}$ \cite{wolsey} and at least $\frac{6}{5}$ \cite{alexander2006integrality}. Therefore, we cannot hope to obtain an LP-based $(\frac{6}{5}-\epsilon)$-approximation algorithm for this problem using this LP relaxation.

@@

One can extend the FDT algorithm for binary IPs into covering $\{0,1,2\}$ IPs by losing a factor $2^{|\spp(x)|}$ on top of the loss for FDT. In order to eradicate this extra factor, we need to treat the coordinate $i$ with $x_i=1$ differently. We focus on the \textsc{2-edge-connected multigraph graph problem (2EC)}: Given a graph $G=(V,E)$ and $c\in \mathbb{R}^{E}_{\geq 0}$ find a 2-edge-connected multi-subgraph (henceforth a multigraph) of $G$ with minimum cost. The natural linear programming relaxation for this problem is 
\begin{equation}
\min \{cx \; : \; x(\delta(U))\geq 2 \; \text{ for }\emptyset \subset U \subset V, \; x\in [0,2]^E\}
\end{equation}
We denote the feasible region of this LP by $\subtour(G)$. Let $\2ec(G)$ be the convex hull of incidence vectors of 2-edge-connected multigraphs of graph $G$. Following the definition in (\ref{gapproblem}) have
\begin{equation}
g(\2ec) = \max_{c\geq 0 , G}\frac{\min_{x\in \2ec(G)} cx}{\min_{x\in \subtour(G)} cx}.
\end{equation}

\begin{restatable}{thm}{FDTEC}
	\label{FDT2EC}
	Let $G=(V,E)$ and $x$ be an extreme point of  $\subtour(G)$. The FDT algorithm for 2EC produces $\lambda\in [0,1]^k$ and 2-edge-connected multigraphs $F_1,\ldots,F_k$ such that $k\leq 2|V|-1$, $\sum_{i=1}^{k}\lambda_i \chi^{F_i}\leq \min(Cx,\{2\}^n)$, and $\sum_{i=1}^{k}\lambda_i = 1$. Moreover, $C\leq g(\2ec)^{|E_x|}$.
\end{restatable}

@@

\subsubsection{2-edge-connected multigraph problem}
Known polyhedral structure makes it easier to study integrality gaps for such problems. We use the idea of fundamental extreme point \cite{carrravi,boydcarr,Carr2004} to create the ``hardest'' LP solutions to decompose.

>>>>

The value of $g(I)$ depends on the constraints in (\ref{S}).  We cannot hope to find solutions for the IP with objective values better than $g(I)\cdot z_{LP}(I,c)$. More generally we can define the integrality gap for a class of instances $\mathcal{I}$ as follows.% In this case, the integrality gap for problem $\mathcal{I}$ is
\begin{equation}\label{gapproblem}
g(\mathcal{I}) = \max_{c\geq 0 , I\in\mathcal{I}}\frac{z_{IP}(I,c)}{z_{LP}(I,c)}.
\end{equation}
For example, the aforementioned integrality gap of the subtour elimination relaxation for the 2EC is at most $\frac{3}{2}$ \cite{wolsey} and at least $\frac{6}{5}$ \cite{alexander2006integrality}. Therefore, we cannot hope to obtain an LP-based $(\frac{6}{5}-\epsilon)$-approximation algorithm for this problem using this LP relaxation.

@@

One can extend the FDT algorithm for binary IPs into covering $\{0,1,2\}$ IPs by losing a factor $2^{|\spp(x)|}$ on top of the loss for FDT. In order to eradicate this extra factor, we need to treat the coordinate $i$ with $x_i=1$ differently. We focus on the \textsc{2-edge-connected multi-subgraph graph problem (2EC)}: Given a graph $G=(V,E)$ and $c\in \mathbb{R}^{E}_{\geq 0}$ find a 2-edge-connected multi-subgraph of $G$ with minimum cost. The natural linear programming relaxation for this problem is 
\begin{equation}
\min \{cx \; : \; x(\delta(U))\geq 2 \; \text{ for }\emptyset \subset U \subset V, \; x\in [0,2]^E\}
\end{equation}
We denote the feasible region of this LP by $\subtour(G)$. Let $\2ec(G)$ be the convex hull of incidence vectors of 2-edge-connected multi-subgraphs of graph $G$. Following the definition in (\ref{gapproblem}) have
\begin{equation}
g(\2ec) = \max_{c\geq 0 , G}\frac{\min_{x\in \2ec(G)} cx}{\min_{x\in \subtour(G)} cx}.
\end{equation}

\begin{restatable}{thm}{FDTEC}
	\label{FDT2EC}
	Let $G=(V,E)$ and $x$ be an extreme point of  $\subtour(G)$. The FDT algorithm for 2EC produces $\lambda\in [0,1]^k$ and 2-edge-connected multi-subgraphs $F_1,\ldots,F_k$ such that $k\leq 2|V|-1$, $\sum_{i=1}^{k}\lambda_i \chi^{F_i}\leq \min(Cx,\{2\}^n)$, and $\sum_{i=1}^{k}\lambda_i = 1$. Moreover, $C\leq g(\2ec)^{|E_x|}$.
\end{restatable}

@@

\subsubsection{2-edge-connected multi-subgraph problem}
Known polyhedral structure makes it easier to study integrality gaps for such problems. We use the idea of fundamental extreme point \cite{carrravi,boydcarr,Carr2004} to create the ``hardest'' LP solutions to decompose.
